---
title: "Covid91 vaccine study Final2024"
author: "Put your name here!"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalSumr2024.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males
$H_0$: The Covid921 vaccine is not effective in males  

$H_A$: The Covid921 vaccine is effective in males   

### Methods 
treatment will be Cat   
sex will be Cat   
helper file we are using is   
CAT~CAT   

### Descriptive Results 

#### Graphical Results 
```{r}
males=filter(FinalData,FinalData$sex=="M")
barchartGC(~treatment + infected,data=males)
barchartGC(~infected + treatment,data=males,type="percent")

```
 
 This chart shows that males who got the actual vaccines are less likely to get Covid921,   
 than those who got the placebo vaccine.       


#### Numerical Results 
```{r}
table1 <- xtabs(~treatment + infected,data=males)
rowPerc(table1)
colPerc(table1)
```
This shows that about 74% males who took the placebo got covid19, where about 26% of males who got the vaccines got covid19 

#### Inferential Results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```
When running the Chi-squared we get a pvalue of .0000000000000002 which is less than 0.05. This means that we reject the null hypothesis that Covid921 vaccine is not effective in males 

#### Conclusion 




## Females

$H_0$: The Covid921 vaccine is not effective in females 

$H_A$: The Covid921 vaccine is effective in females

### Methods 

### Descriptive Results 

#### Graphical Results 

#### Inferential Results 

#### Conclusion 




## LGBTQ
$H_0$: The Covid921 vaccine is not effective in LGBTQ Community 

$H_A$: The Covid921 vaccine is effective in LGBTQ  Community 

### Methods 

### Descriptive Results 

#### Graphical Results 

```{r}

LGBTQ=filter(FinalData,FinalData$LGBTQ=="s")
barchartGC(~treatment + infected,data=strai)
barchartGC(~infected + treatment,data=FinalData,type="percent")

```

#### Numerical Results 

#### Inferential Results 

#### Conclusion 



## DrugUser

$H_0$: The Covid921 vaccine is not effective in Drug users 

$H_A$: The Covid921 vaccine is effective in Drug users  

### Methods 

### Descriptive Results 

#### Graphical Results 

#### Numerical Results 

#### Inferential Results 

#### Conclusion 




