---
title: "Covid91 vaccine study Final2024"
author: "Put your name here!"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalSumr2024.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males
$H_0$: The Covid921 vaccine is not effective in males  

$H_A$: The Covid921 vaccine is effective in males   

### Methods 
treatment will be Cat   
infected will be Cat   
helper file we are using is   
CAT~CAT   

### Descriptive Results 

#### Graphical Results 
```{r}
males=filter(FinalData,FinalData$sex=="M")
barchartGC(~treatment + infected,data=males)
barchartGC(~infected + treatment,data=males,type="percent")

```
 
 This chart shows that males who got the actual vaccines are less likely to get Covid921,   
 than those who got the placebo vaccine.  Out of those who got placebo and the vaccine about 50% where ok. 


#### Numerical Results 
```{r}
table1 <- xtabs(~treatment + infected,data=males)
rowPerc(table1)
colPerc(table1)
```
This shows that about 74% males who took the placebo got covid19, where about 26% of males who got the vaccines got covid19 

#### Inferential Results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```
When running the Chi-squared we get a pvalue of .0000000000000002 which is less than 0.05. This means that we reject the null hypothesis that Covid921 vaccine is not effective in males 

When running the Fisher Exact test we have .0000000000000002 which is also less than 0.05. 

The odds ratio would be 2.81 : 1, which means that their is a 2.81 is more likely to get when taking placebo, than those who are 1 times more likely to get covid  when taking the vaccines.   

#### Conclusion 

Our question was to find if the Covid19 is effected in males or not effective in males. When doing our graphical result it shows that males who got the covid vaccines are less likely to get covid then those who got Placebo. When indicating the numerical results, it shows that about 74% males who took the placebo got covid19, where about 26% of males who got the vaccines got covid19. We ran the Chi-square and got the Pvalue of .0000000000000002 which is less than .05. This indicates that we reject the null hypothesis. When doing the fisher test we came out with the same hypothesis as the chi-sqr. There for we reject the null hypothesis again. We conclude that taking the covid19 vaccine is effective in males. 




## Females

$H_0$: The Covid921 vaccine is not effective in females 

$H_A$: The Covid921 vaccine is effective in females

### Methods 
treatment will be Cat   
infected will be Cat   
helper file we are using is   
CAT~CAT 
### Descriptive Results 

#### Graphical Results 

```{r}
females=filter(FinalData,FinalData$sex=="F")
barchartGC(~treatment + infected,data=females)
barchartGC(~infected + treatment,data=females,type="percent")

```

This chart shows that females who got the actual vaccines are less likely to get Covid921,   
 than those who got the placebo vaccine.  Out of those females who got placebo and are vaccinated about 50% where okay. 
 
#### Numerical Results 

```{r}
table1 <- xtabs(~treatment + infected,data=females)
rowPerc(table1)
colPerc(table1)
```
 When doing the numerical results it shows that out of the 100% about 65% of women who got placebos are more likely to get Covid19 than those who actually got the vaccines which was about 35% of women.  
 
 Out of the 100% of both vaccinated and placebo females both were about 50/50 okay. 
 
#### Inferential Results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

When running the chi- sqr it states that the Pvalue is 1.914e-08. This will calculate to .00000001914. This means that we reject our null hypothesis because it was less than 0.05.This mean that there is a difference in women who use vaccines.   

When running the fisher Exact test it shows that the Pvalue is also smaller than 0.05. Our Pvalue comes out to .00000001246, which is also smaller than 0.05.    

When calculating the odds ratio 1.86:1 is more likely to get covid when taking placebo, than those who are 1 times more likely to get covid when taking the vaccines.

#### Conclusion 
 Our objective was to check if the vaccine was effective with females or not effective with females. When running the graph data that shows it showed that females who where on placebo where more likely to get covid 19 than those who got the vaccines. Out of all the females who got the vaccine and the placebo about 50% of them were okay. We ran our numerical charts which shows that 65% of females on placebo are more likely to get covid then 35% of the females who got the vaccinated. When running our chi square and fisher test we concluded that because our Pvalue was smaller than 0.05 we reject the null hypothesis that Covid921 vaccine is not effective in females. Inconclusion The Covid921 vaccines is effective in females.  

## LGBTQ
$H_0$: The Covid921 vaccine is not effective in LGBTQ Community 

$H_A$: The Covid921 vaccine is effective in LGBTQ  Community 

```{r}
LGBTQ = filter(FinalData, LGBTQ == "gay")
```


### Methods 

treatment will be Cat   
infected will be Cat   
helper file we are using is   
CAT~CAT 

### Descriptive Results 

#### Graphical Results 


```{r}
barchartGC(~infected + treatment,data=LGBTQ)
barchartGC(~infected + treatment,data=LGBTQ, type="percent")

```



#### Numerical Results 


```{r}
table1 <- xtabs(~infected + treatment, data=LGBTQ)
rowPerc(table1)
colPerc(table1)
```


#### Inferential Results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```


#### Conclusion 



## DrugUser

$H_0$: The Covid921 vaccine is not effective in Drug users 

$H_A$: The Covid921 vaccine is effective in Drug users  

```{r}
DrugUser = filter(FinalData, DrugUser == "yes")
```


### Methods 

treatment will be Cat   
infected will be Cat   
helper file we are using is   
CAT~CAT 

### Descriptive Results 

#### Graphical Results 

```{r}
DrugUser=filter(FinalData,FinalData $DrugUser =="yes")
barchartGC(~ infected+treatment ,data=DrugUser)
barchartGC(~treatment + infected  ,data=DrugUser,type="percent")

```



#### Numerical Results 

```{r}
table1 <- xtabs(~infected + treatment, data=DrugUser)
rowPerc(table1)
colPerc(table1)
```



#### Inferential Results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```





#### Conclusion 




