---
title: "Covid91 vaccine study Final2024"
author: "Sandra Iovino !"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date:  8/1/2024 "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalSumr2024.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males
$H_0$: The Covid921 vaccine is not effective in males  

$H_A$: The Covid921 vaccine is effective in males   

### Methods 
treatment will be Cat   
infected will be Cat   
helper file we are using is   
CAT~CAT   

### Descriptive Results 

#### Graphical Results 
```{r}
males=filter(FinalData,FinalData$sex=="M")
barchartGC(~treatment + infected,data=males)
barchartGC(~infected + treatment,data=males,type="percent")

```
 
 This chart shows that males who got the actual vaccines are less likely to get Covid921,   
 than those who got the placebo vaccine.  Out of those who got placebo and the vaccine about 50% were OK. 


#### Numerical Results 
```{r}
table1 <- xtabs(~treatment + infected,data=males)
rowPerc(table1)
colPerc(table1)
```
This shows that about 74% males who took the placebo got covid19, where about 26% of males who got the covid vaccines got covid19. Out of the 100% that were okay about 50/50 were placebo and vaccines 

#### Inferential Results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```
When running the Chi-squared we gott a pvalue of .0000000000000002 which is less than 0.05. This means that we reject the null hypothesis that Covid921 vaccine is not effective in males. We conclude that Covid921 vaccine is effective in males.

When running the Fisher Exact test we have .0000000000000002 which is also less than 0.05. 

The odds ratio would be 2.81 : 1, which means that their is a 2.81 times more likely to get covid19 when taking placebo, than those who are 1 times more likely to get covid19 when taking the vaccines.   

#### Conclusion 

Our question was to find if the Covid19 is effected in males or not effective in males. When doing our graphical result it shows that males who got the covid19 vaccines are less likely to get covid19 then those who got Placebo. When indicating the numerical results, it shows that about 74% males who took the placebo got covid19, where about 26% of males who got the vaccines got covid19. We ran the Chi-square and  got the Pvalue of .0000000000000002 which is less than 0.05. This indicates that we reject the null hypothesis. When doing the fisher test we came out with the same hypothesis as the chi-sqr. There for we reject the null hypothesis again. We conclude that taking the covid19 vaccine is effective in males. 

## Females

$H_0$: The Covid921 vaccine is not effective in females 

$H_A$: The Covid921 vaccine is effective in females

### Methods 

treatment will be Cat   
infected will be Cat   
helper file we are using is   
CAT~CAT 
### Descriptive Results 

#### Graphical Results 

```{r}
females=filter(FinalData,FinalData$sex=="F")
barchartGC(~treatment + infected,data=females)
barchartGC(~infected + treatment,data=females,type="percent")

```

This chart shows that females who got the actual vaccines are less likely to get Covid921,   
 than those who got the placebo vaccine.  Out of those females who got placebo and are vaccinated about 50/50 where okay. 
 
#### Numerical Results 

```{r}
table1 <- xtabs(~treatment + infected,data=females)
rowPerc(table1)
colPerc(table1)
```
 When doing the numerical results it shows that out of the 100% about 65% of women who got placebos got  Covid19 than those who actually got the vaccines which was about 35% of women.  
 
#### Inferential Results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

When running the chi- sqr it states that the Pvalue is 1.914e-08. This will calculate to .00000001. This means that we reject our null hypothesis because it was less than 0.05. We conclude that there is a difference in women who use vaccines.   

When running the fisher Exact test it shows that the Pvalue is also smaller than 0.05. Our Pvalue comes out to .00000001, which is also smaller than 0.05.    

When calculating the odds ratio 1.86:1 . 
1.86 times more likely to get covid19 when taking placebo, than those who are 1 times more likely to get covid when taking the covid19 vaccines.

#### Conclusion 
 Our objective was to check if the vaccine was effective with females or not effective with females. When running the graphical data, it shows that females who where on placebo where more likely to get covid 19 than those who got the vaccines. We ran our numerical results, which shows thatout of the 100% about 65% of females on placebo got covid19 and 35% of the females who got the vaccine got covid19. When running our chi square and fisher test we concluded that because our Pvalue was smaller than 0.05 we reject the null hypothesis that Covid921 vaccine is not effective in females. In conclusion The Covid921 vaccines is effective in females.  

## LGBTQ
$H_0$: The Covid921 vaccine is not effective in LGBTQ Community 

$H_A$: The Covid921 vaccine is effective in LGBTQ  Community 

```{r}
LGBTQ = filter(FinalData, LGBTQ == "gay")
```


### Methods 

treatment will be Cat   
infected will be Cat   
helper file we are using is   
CAT~CAT 

### Descriptive Results 

#### Graphical Results 


```{r}
barchartGC(~infected + treatment,data=LGBTQ)
barchartGC(~infected + treatment,data=LGBTQ, type="percent")

```
 
 When running this bar graph it shows that participant who are LGBTQ and are on the placebo are more likely to get covid19 then those who took the vaccines covid19. Out of both placebo and vaccine  it looks like about 47% of placebo were okay and about 50% of vaccines were okay.    


#### Numerical Results 


```{r}
table1 <- xtabs(~ treatment+infected , data=LGBTQ)
rowPerc(table1)
colPerc(table1)
```
When doing the numerical results it shows that out of the 100% about 80% of LGBTQ people who got placebos are more likely to get Covid921 than those who actually got the vaccines which was about 20% of LGTBQ.  
 
 Out of the 100% about 48.33% that were placebo were okay and 51.67% that were vaccines were OK.   
 
#### Inferential Results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```
When running the chi- sqr it states that the Pvalue is 2.266e-06. This will calculate to .000002266. This means that we reject our null hypothesis because it was less than 0.05.This mean that there is a difference in the LGBTQ community who use vaccines.   

When running the fisher Exact test it shows that the Pvalue is also smaller than 0.05. Our Pvalue comes out to .00000007269, therefor we reject the null hypothesis. 
When calculating the odds ratio 4.27:1, 4.27 times is more likely to get covid when taking placebo, than those who are 1 times more likely to get covid when taking the vaccines.


#### Conclusion 

Our job was to find if the $H_0$: The Covid921 vaccine is not effective in LGBTQ Community. When running the graphical results it shows that participant who are in the LGBTQ community are most likely to get covid19 than those who got the covid19 vaccine. When running the numerical results it shows that  out od the 100% about 80% the LGBTQ community who are on placebo are most likely to get covid19 then those 20% of participates who are LGBTQ who got the covid19 vaccine. When running the P value for both fisher and chi- Square, we conclude that both Pvalues are smaller than 0.05. This mean that we reject our null hypothesis and conclude that Covid921 vaccine is effective in the LGBTQ community.  

## DrugUser

$H_0$: The Covid921 vaccine is not effective in Drug users 

$H_A$: The Covid921 vaccine is effective in Drug users  

```{r}
DrugUser = filter(FinalData, DrugUser == "yes")
```


### Methods 

treatment will be Cat   
infected will be Cat   
helper file we are using is   
CAT~CAT 

### Descriptive Results 

#### Graphical Results 
```{r}
DrugUser=filter(FinalData,FinalData$DrugUser=="yes")
barchartGC(~treatment + infected,data=DrugUser)
barchartGC(~infected + treatment,data=DrugUser,type="percent")

```

  When running graphical results it shows that 97% of the DrugUsers who took the covid19 Vaccine are most likely to get covid 19 than those who took the placebo which would be about 10% of the DrugUsers.   


#### Numerical Results 

```{r}
table1 <- xtabs(~treatment +infected , data=DrugUser)
rowPerc(table1)
colPerc(table1)
```
The Numerical Results shows that out of the 100%, 92% of the DrugUsers got more covid19 than the 8%  who where on the placebo.


#### Inferential Results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

When running the Chi-squared we get a pvalue of .0000000001,  which is less than 0.05. This means that we reject the null hypothesis that Covid921 vaccine is not effective in DrugUsers.

When running the Fisher Exact test we have a Pvalue of .000000000003 which is also less than 0.05. 

The odds ratio would be 0.9 : 1, which means that their is a 0.9 times more likely to get covid19 when taking placebo, than those who are 1 times more likely to get covid  when taking the vaccines.   


#### Conclusion 

Our job was to find if the $H_0$: The Covid921 vaccine is not effective in DrugUsers. When running the graphical results it shows that DrugUsers who got the covid 19 vaccines are more likely to get covid19 then those who took the Placebo. Now, when running the numerical results, it also stated that 92% of DrguUsers who got the covid19 vaccine are more likely to get covid19. When seeing the placebo data it show that about only 8% of DrugUsers got covid 19. When Running the chia and fisher test we got a Pvalue that is smaller than 0.05, so we reject our null Hypothesis Covid921 vaccine is not effective in DrugUsers. Our final conclusion shows that Covid921 vaccine is effective in DrugUsers. It looks like the DrugUsers who got the covid vaccines are actually getting more sick then those who got the placebo. 



